---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# harpTobac

<!-- badges: start -->
<!-- badges: end -->

_harpTobac_ is a an interface to the Python package, 
[tobac](https://tobac.readthedocs.io/en/latest/index.html), for tracking and 
object based analysis of clouds. 

It currently enables you to pass `harp_grid_df` data frames to do a basic 
analysis and tracking of 2d fields. That is, feature detection in 2d, 
segmentation in 2d and 2d linking of features into tracks. 

Currently no specific functions for visualisation have been written, but the 
outputs are basic data frames and `harp_grid_df` data frames that can be 
easily visualised using [ggplot2](https://ggplot2.tidyverse.org/) and 
[harpVis](https://harphub.github.io/harpVis/reference/index.html) geoms. 

## Installation

You can install harpTobac from Github like so:

```{r eval=FALSE}
if (!requireNamespace("remotes")) {
  install.packages("remotes")
}
remotes::install_github("harphub/harpTobac")
```

Once you have installed _harpTobac_ you will need to install the _tobac_ Python 
package. This is done using the `install_tobac()` function. The necessary 
packages will be installed into a _virtualenv_ called "harp-py". This is where 
all Python packages that can be used by _harp_ are installed and run from. If 
you already have a "harp-py" _virtualenv_ you will be asked if you want to 
remove it prior to installation. Typically you will answer no to this, otherwise
all of your harp Python packages will be removed. 

```{r eval=FALSE}
library(harpTobac)
install_tobac()
```

You will only need to do the above once, straight after installing _harpTobac_. 
Subsequently, when the _harpTobac_ package is attached your session will use the 
"harp-py" _virtualenv_. 

