---
title: "Track satellite brightness temperature"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Track satellite brightness temperature}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The purpose of this article is to show how to read in data that harp cannot
natively deal with. In this case we will read in SEVIRI brightness temperature 
data from a NetCDF file that lacks information about the domain. 

```{r setup, results='hide', message=FALSE}
library(harpTobac)
library(harpCore)
library(harpVis)
library(ncdf4)
```

# Attach the data file and define the domain

The data are stored as example data with the _harpTobac_ package and can be 
accessed via the `system.file()` function. 

```{r attach-nc}
file_name <- system.file(
  "example_data/Spain_202205_SEVIRI_bt_all.nc",
  package = "harpTobac"
)
nc <- nc_open(file_name)
```

Longitude and latitude are stored as 2d arrays, so let's read them in and then 
close the connection to the NetCDF file.

```{r read-lon-lat}
lon <- ncvar_get(nc, "longitude")
lat <- ncvar_get(nc, "latitude")
nc_close(nc)
```


Now we need to establish the grid resolution. We can establish if the data are
on a regular grid, by checking if the mean difference between each value is the 
same as the maximum. 

```{r check-for-regular-grid}
identical(apply(diff(lon), 1, mean), apply(diff(lon), 1, max))
identical(apply(diff(t(lat)), 1, mean), apply(diff(t(lat)), 1, max))
```

Now that we have established that the data are on a regular grid, we can 
extract the horizontal resolution. 

```{r get-dx-dy}
dx <- diff(lon)[1]
dy <- diff(t(lat))[1]
```

Finally we need the centre point of the domain in order to properly define 
the domain.

```{r domain-centre}
centre_lon <- min(lon) + diff(range(lon)) / 2
centre_lat <- min(lat) + diff(range(lat)) / 2
```

Now we can define the domain.

```{r define-domain, fig.width=8, fig.height=8, fig.align='center'}
dom <- define_domain(
  centre_lon = centre_lon, 
  centre_lat = centre_lat,
  nxny       = dim(lon),
  dxdy       = c(dx, dy),
  proj       = "longlat"
)

plot(dom)
```

We also need to get the time data from the file. These are stored in hours 
since the first date-time, but we want the data to be the actual date times. 
_harpIO_ has a, currently non exported function, that can read the time data 
and convert them in date-times in UTC. 

```{r read-date-times}
valid_dttm <- unixtime_to_dttm(
  harpIO:::get_time_nc(file_name, harpIO::netcdf_opts())$validdate
)
```

# Read the brightness temperature data

The brightness temperature is stored in degrees C, which we will convert to 
Kelvin. We will read in one time at a time and use the domain e have just 
defined to put the data into a geofield, and subsequently a geolist. We will 
do this with the help of `lapply()` and put the data in a tibble (just a 
data frame with a nicer print method). 

```{r read-brightness-temp}
library(tibble)
nc <- nc_open(file_name)
bt <- tibble(
  valid_dttm = valid_dttm,
  brightness_temp = geolist(
    lapply(
      seq_along(valid_dttm),
      function(i) geofield(
        ncvar_get(nc, "bt", start = c(1, 1, i), count = c(-1, -1, 1)) + 273.15,
        domain = dom
      )
    )
  )
)
nc_close(nc)
```

Finally the _harpTobac_ functions require the data to be of class `harp_df` and 
we can easily do that conversion with `as_harp_df()`

```{r to-harp-df}
bt <- as_harp_df(bt)
```

# Feature detection

We will detect features using minimum brightness temperature thresholds of 250, 
240, 230 and 220 K.

```{r detect-features, results='hide'}
features <- detect_features_multithreshold(
  bt, 
  thresholds         = seq(250, 220, -10),
  data_col           = brightness_temp,
  target             = "min",
  n_min_threshold    = 16, 
  position_threshold = "weighted_diff"
)
```

And plot the feature locations at each time.

```{r plot-features, fig.width=8, fig.height=8, fig.align='center'}
countries <- get_map(get_domain(bt$brightness_temp), polygon = FALSE)
ggplot(features, aes(longitude, latitude)) +
  geom_path(aes(x, y), countries, colour = "grey30") + 
  geom_point(aes(colour = factor(threshold_value))) +
  facet_wrap(~format(as.POSIXct(timestr, tz = "UTC"), "%H:%M")) +
  coord_equal(expand = FALSE) + 
  theme_harp_map() + 
  labs(colour = bquote(atop(Brightness~Temp, Threshold~"["*K*"]")))
```

# Segmentation

We will segment the data using a threshold of 250K.

```{r segmentation, results='hide'}
library(zeallot)
c(segments, features) %<-% segment_2d(
  features, 
  bt, 
  threshold = 250, 
  data_col  = brightness_temp, 
  target    = "min"
)
```
 
And plot with the brightness temperature for some selected times.

```{r plot-segments, fig.width=8, fig.height=8, fig.align='center'}
ggplot() + 
  geom_georaster(
    aes(geofield = brightness_temp), bt[13:16, ], 
    upscale_factor = 4, upscale_method = "downsample"
  ) +
  geom_geocontour(
    aes(geofield = segmentation_mask), segments[13:16, ],
    colour = "red"
  ) + 
  geom_path(aes(x, y), countries, colour = "grey30") +
  facet_wrap(~valid_dttm) +
  scale_fill_viridis_c(direction = -1) +
  coord_equal(expand = FALSE) +
  theme_harp_map()
```

# Cell tracks

Now we can compute the cell tracks. 

```{r compute-tracks, results='hide'}
tracks <- link_tracks(
  features, 
  bt, 
  data_col        = brightness_temp, 
  v_max           = 20, 
  stubs           = 2,  
  subnetwork_size = 100, 
  adaptive_step   = 0.95, 
  adaptive_stop   = 0.2
)
```

And plot them, separately for each feature threshold.

```{r plot-tracks, fig.width=8, fig.height=8, fig.align='center'}
countries <- get_map(get_domain(bt$brightness_temp))
ggplot(filter(tracks, cell > -1), aes(x, y)) + 
  geom_polygon(aes(group = group), countries, fill = "grey", colour = "grey30") +
  geom_path(
    aes(group = factor(cell), colour = factor(threshold_value)),
    arrow = arrow(type = "open", angle = 30, length = unit(0.1, "cm"))
  ) +
  facet_wrap(~paste0("Brightness temp >= ", threshold_value, "K")) +
  labs(colour = bquote(atop(OLR~threshold, "["*W.m^{-2}*"]"))) +
  coord_equal(expand = FALSE) + 
  theme_harp_map()
```

