---
title: "Track OLR features"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tracking_model_OLR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

In this article, we show an example of tracking features in an outgoing longwave
radiation field from the NWP model MEPS. We are going to make use of the 
_harpIO_ package to load the data and the _ggplot2_ package, together with some
expansions from the _harpVis_ package to plot the data.  

```{r setup, results='hide', message=FALSE}
library(harpIO)
library(harpVis)
library(harpTobac)
```

```{r echo=FALSE}
if (!reticulate::virtualenv_exists("harp-py")) {
  install_tobac()
}
```

# Loading the data

First we need to load the data using `read_forecast()` from the _harpIO_ 
package. The data are downloaded from the open data Thredds server at the 
Norwegian Meteorological Institute. 


There are a few steps we need to go through to ensure we read the data 
properly. First we'll add the outgoing longwave parameter to the parameter 
definitions so that `read_forecast()` knows what parameter to extract. The data 
are also negative for outgoing radiation so we multiply by -1

```{r param-defs}
my_params <- add_param_def(
  "olr", 
  netcdf = new_netcdf_param("toa_outgoing_longwave_flux"), 
  func = function(x) x * -1
)
```

Then we set the path and template for the file(s) to read, as well as some 
information about the netcdf files. 
```{r path-template}
url       <- "https://thredds.met.no/thredds/dodsC/meps25epsarchive"
template  <- "{YYYY}/{MM}/{DD}/{fcst_model}_sfc_{YYYY}{MM}{DD}T{HH}Z.ncml"
file_opts <- netcdf_opts(
  z_var = "top_of_atmosphere", ref_time_var = "forecast_reference_time"
)
```

And now we can read the data...
```{r read-data}
olr <- read_forecast(
  dttm             = 2024052700,
  fcst_model       = "meps_det",
  parameter        = "olr",
  lead_time        = seq(0, 23),
  file_path        = url, 
  file_template    = template, 
  file_format      = "netcdf",
  file_format_opts = file_opts,
  param_defs       = my_params, 
  return_data      = TRUE 
)
```

# Feature detction

Feature detection is done using thresholds to identify contiguous features in 
the field of interest at each time. This requires careful selection of 
thresholds, so as a first step we will plot the OLR at each time step. 

```{r olr-plot, fig.width=8, fig.height=8, fig.align='center'}
countries <- get_map(get_domain(olr$fcst), polygon = FALSE)
ggplot() + 
  geom_georaster(
    aes(geofield = fcst), olr, 
    upscale_factor = 8, upscale_method = "downsample"
  ) + 
  geom_path(aes(x, y), countries, colour = "grey30") +
  facet_wrap(~format(valid_dttm, "%H:%M")) +
  scale_fill_viridis_c(bquote(OLR~"["*W.m^{-2}*"]"), direction = -1) + 
  coord_equal(expand = FALSE) +
  theme_harp_map()
```

It looks like most of the interesting features are for OLR < ~180. We can get a 
better idea by masking all values that are > 180, by adding limits to the 
colour scale. 

```{r olr-mask-plot, fig.width=8, fig.height=8, fig.align='center'}
countries <- get_map(get_domain(olr$fcst), polygon = FALSE)
ggplot() + 
  geom_georaster(
    aes(geofield = fcst), olr, 
    upscale_factor = 8, upscale_method = "downsample"
  ) + 
  geom_path(aes(x, y), countries, colour = "grey30") +
  facet_wrap(~format(valid_dttm, "%H:%M")) +
  scale_fill_viridis_c(
    bquote(OLR~"["*W.m^{-2}*"]"), direction = -1, 
    limits = c(NA, 180), na.value = "white"
  ) + 
  coord_equal(expand = FALSE) +
  theme_harp_map()
```

Using this information we will use thresholds of 180, 160 and 140 in the 
feature detection, setting the target to "min" since we want features to be 
identified that are less than the thresholds. Features are detected using 
`detect_features_multithreshold()` that is a wrapper around [tobac.feature_detection.feature_detection_multithreshold](https://tobac.readthedocs.io/en/latest/tobac.html#module-tobac.feature_detection)
from the original Python package. We will only detect features that are at least 
16 contiguous pixels in size, and use the "weighted_diff" method to set the 
position of the features. 

```{r feature-detection, results='hide'}
features <- detect_features_multithreshold(
  olr, 
  thresholds         = c(180, 160, 140),
  data_col           = fcst,
  target             = "min",
  n_min_threshold    = 16, 
  position_threshold = "weighted_diff"
)
```

We can now plot the feature locations for each threshold and time.

```{r plot-features, fig.width=8, fig.height=8, fig.align='center'}
ggplot(features, aes(projection_x_coordinate, projection_y_coordinate)) +
  geom_path(aes(x, y), countries, colour = "grey30") + 
  geom_point(aes(colour = factor(threshold_value))) +
  facet_wrap(~format(as.POSIXct(timestr, tz = "UTC"), "%H:%M")) +
  coord_equal(expand = FALSE) + 
  theme_harp_map() + 
  labs(colour = bquote(atop(OLR~Threshold,"["*W.m^{-2}*"]")))
```

# Segmentation

Having run the feature detection, the next step is to associate regions with 
the identified features. This is done using `segment_2d()` that is a wrapper 
around the [tobac.segmentation.segmentation_2D](https://tobac.readthedocs.io/en/latest/tobac.html#module-tobac.segmentation)
function from the original Python package. Here a single threshold is needed, 
and we use the highest of the thresholds we used in the feature detection. 
`segment_2d` returns a named list, but we can destructure using the `%<-%` 
operator from the _zeallot_ package.

```{r segmentation, results='hide'}
library(zeallot)
c(segments, features) %<-% segment_2d(
  features, 
  olr, 
  threshold = 180, 
  data_col  = fcst, 
  target    = "min"
)
```

`segments` is a data frame with a _geolist_ coloumn that contains fields with a 
mask identifying the areas that are associated with each feature, while 
`features` is the same as the output of the previous step, but with an extra 
column giving the number of cells associated with each feature. The segment 
masks can be overlayed on top of the original OLR data. Here we just show 4 
times. 

```{r plot-segment-mask, fig.width=8, fig.height=8, fig.align='center'}
ggplot() + 
  geom_georaster(
    aes(geofield = fcst), olr[11:14, ], 
    upscale_factor = 4, upscale_method = "downsample"
  ) +
  geom_geocontour(
    aes(geofield = segmentation_mask), segments[11:14, ],
    colour = "red"
  ) + 
  geom_path(aes(x, y), countries, colour = "grey30") +
  facet_wrap(~valid_dttm) +
  scale_fill_viridis_c(bquote(atop(OLR,"["*W.m^{-2}*"]")), direction = -1) +
  coord_equal(expand = FALSE) +
  theme_harp_map()
```

# Tracking

With the features identified, the tracking algorithm attempts to link these 
features in time to produce tracks. The tracking is computed with the 
`link_tracks()` function that is a wrapper around 
[tobac.tracking.linking_trackpy](https://tobac.readthedocs.io/en/latest/tobac.html#module-tobac.tracking)
from the original Python package. 

```{r link-tracks, results='hide'}
tracks <- link_tracks(
  features, 
  olr, 
  data_col        = fcst, 
  v_max           = 20, 
  stubs           = 2,  
  subnetwork_size = 100, 
  adaptive_step   = 0.95, 
  adaptive_stop   = 0.2
)
```

Some of the arguments in the call to `link_tracks()` require some 
experimentation, and are better explained in some of the [tobac examples](https://tobac.readthedocs.io/en/latest/examples.html).

The tracks for each of the identified cells can easily be plotted. For all 
tracks throughout the period this can be done making sure to remove untracked 
cells (these are labelled -1).

```{r plot-tracks, fig.width=8, fig.height=8, fig.align='center', message=FALSE}
library(dplyr)
countries <- get_map(get_domain(olr$fcst))
ggplot(filter(tracks, cell > -1), aes(x, y)) + 
  geom_polygon(aes(group = group), countries, fill = "grey", colour = "grey30") +
  geom_path(
    aes(group = factor(cell), colour = factor(threshold_value)),
    arrow = arrow(type = "open", angle = 30, length = unit(0.1, "cm"))
  ) +
  labs(colour = bquote(atop(OLR~threshold, "["*W.m^{-2}*"]"))) +
  coord_equal(expand = FALSE) + 
  theme_harp_map()
```

We could then plot some statistics, such as the distribution of the lifetimes of
individual cells.

```{r plot-cell-lifetime, fig.width=6, fig.align='center'}
ggplot(
  summarise(
    filter(tracks, cell > -1),
    lifetime = max(time_cell) / 3600, 
    .by      = cell
  )
) +
  geom_bar(aes(x = factor(lifetime)), fill = "steelblue") + 
  labs(x = "Cell lifetime [h]")
  
```

# Epilogue

There are many analysis possibilities with these data. In addition the 
_harpTobac_ package gives you access to all of the functionality of the Python
tobac package via `tobac$<module name>$<function>`.

